What paradigms can webcam eye-tracking be used for? Attempted replications of 5 ``classic'' cognitive science experiments

### 2023-08-17 manuscript.Rmd

The use of eye-tracking to study cognition took off when Alfred Yarbus used suction cups to affix a mirror system to the sclera of the eye in order to monitor eye position during the perception of images [@yarbus1967eye]. In one study, participants viewed a painting depicting multiple people in a complex interaction inside of a 19th century Russian home. Yarbus showed, among other things, that the scan paths and locations of fixations were largely dependent on the instructions given to participants (e.g., View the picture freely vs. Remember the position of the people and objects in the room). In other words, the cognitive processing that the individual is engaged in drives the visuo-motor system. Since these findings, eye-tracking has become a central method in cognitive science research [for reviews, see @hayhoe2005eye, @richardsonEyeTrackingResearch2004, @rayner1998eye]. For example, gaze location during natural scene perception is used to test theories of visual attention [e.g., @henderson2017meaning]. And eye-movements during auditory language comprehension demonstrated the context-dependent and incremental nature of language processing [e.g., @tanenhaus1995].

@@

A few previous studies have used webcam eye-tracking in the context of behavioral/cognitive science experiments and reported on the quality of the data. In what follows, we provide a brief review of the published work that we are aware of (note that we focus on studies with adult participants as the considerations for eye-tracking of children are substantially different, but see e.g., @banki2022comparing).

@semmelmann2018online compared eye-tracking results between in-lab and online studies in three tasks: fixation, pursuit, and free viewing.

Yang & Krajbich,

slim & hartsuiker

>>>>

The use of eye-tracking to study cognition took off when Alfred Yarbus used suction cups to affix a mirror system to the sclera of the eye in order to monitor eye position during the perception of images [@yarbus1967eye]. In one study, participants viewed a painting depicting multiple people in a complex interaction inside of a 19th century Russian home. Yarbus showed, among other things, that the scan paths and locations of fixations were largely dependent on the instructions given to participants (e.g., View the picture freely vs. Remember the position of the people and objects in the room). In other words, the cognitive processing that the individual is engaged in drives the visuo-motor system. Since these findings, eye-tracking has become a central method in cognitive science research [for reviews, see @hayhoe2005eye, @richardsonEyeTrackingResearch2004, @rayner1998eye]. For example, gaze location during natural scene perception is used to test theories of visual attention [e.g., @henderson2017meaning]. And eye-movements during auditory language comprehension, using the "visual world paradigm," demonstrated the context-dependent and incremental nature of language processing [e.g., @tanenhaus1995].

@@

A few previous studies have used webcam eye-tracking in the context of behavioral/cognitive science experiments and reported on the quality of the data. In what follows, we provide a brief review of the published work that we are aware of [note that we focus on studies with adult participants as the considerations for eye-tracking of children are substantially different, but see e.g., @banki2022comparing].

@semmelmann2018online compared eye-tracking results between in-lab (n=29) and online (n=28) studies in three tasks: fixation, pursuit, and free viewing. They used the Webgazer.js library [@papoutsaki2016webgazer] and programmed tasks in HTML/Javascript directly. The first two tasks tested measurement of basic gaze properties. In the fixation task, participants were asked to fixate a dot, and in the pursuit task, they were asked to follow a dot with their eyes. In the free viewing task, the eye movements may have been more semantically driven: participants were shown a picture of a human face and asked to look wherever they wanted to on the image. Webgazer was successful in accurately detecting fixation locations and saccades in all three tasks, though online data (collected through a crowdsourcing platform) were slightly more variable and delayed. The free viewing task replicated previously observed statistical patterns (e.g., more fixations on eyes than mouths).

Similarly, @yang2021webcam used the Webgazer.js library [@papoutsaki2016webgazer] combined with the jsPsych library for conducting behavioral experiments in a web browser [@deleeuwJsPsychJavaScriptLibrary2015], to replicate a well-established link between value-based decision-making and eye gaze [e.g., @shimojoGazeBiasBoth2003, @krajbich2010visual]. Online participants (n=38) first saw a series of images of 70 snack foods and rated how much they liked each one. During the primary task, on each of 100 trials, two of the snack food images were displayed on the left and right sides of the display and participants chose the one that they preferred while their gaze was monitored. As in previous work, participants were biased to choose the option they had spent more time looking at. The authors also implemented a code modification to address temporal delays in WebGazer.

Finally, @slim2022moving used the Webgazer.js library via the PCIbex online experiment platform [@zehr2018PCIbex] to replicate a visual world paradigm study in which participants (n=90) listened to sentences (e.g., Mary reads a letter)  while viewing four pictures displayed across the four quadrants of the screen (e.g., a letter, a backpack, a car and a wheelchair)  [@dijkgraaf2017predicting]. When the verb in the sentence was _constraining_ (e.g., read) with respect to the display (only the letter would be an appropriate continuation), listeners made more fixations to the target image (letter) than when the verb was _neutral_ (e.g., Mary steals a letter), as in previous work. However, they observed a substantial delay of ~200ms in the onset of the effect relative to the in-lab study with an infrared eyetracker and a reduction in effect size (despite a threefold increase in sample size relative to the original). The authors noted that delay was particularly problematic given that the purpose of the original study was to capture aspects of predictive or anticipatory processing: in the original study, but not in the web replication, the difference between _constraining_ and _neutral_ conditions emerged before the onset of the final noun (e.g., letter).
